GxHash: A High-Throughput, Non-Cryptographic Hashing Algorithm Leveraging Modern CPU Capabilities

### 2023-09-19 article/article.tex
Update article.tex

\documentclass[11pt]{article} 
\usepackage[margin=0.8in]{geometry} % Sets all margins to 1 inch
\usepackage{amsmath}
\usepackage{hyperref} 
\usepackage{listings}
\usepackage{xcolor}
\usepackage{multicol}

\input{rust-listings.tex}  % Include the language definition

@@

\subsection{Merkle–Damgård Construction}

@@

\subsection{Wide-Pipe Construction} \label{widepipe}

@@

While compilers and CPUs employ various techniques to optimize ILP, their capabilities are often constrained by the inherent data dependencies in the code. It's for this reason that algorithmic design can be pivotal. By structuring algorithms to minimize dependency chains from the outset, we create opportunities for higher ILP that even the most advanced compiler optimizations and CPU features cannot achieve alone. Therefore, algorithmic design that is mindful of ILP can be a game-changer for performance optimization.
\\\\
\textbf{Example}\\
Let's take a FNV-like hashing function. The "naive" way to process an array of elements would look like the \texttt{baseline} method, as shown in Rust snippet (Figure~\ref{fig:ilp_rust_example}). As we can see, every loop iteration requires the value of \texttt{h}, which has been computed the iteration before. Here we have a dependency chain, preventing the compiler from doing any optimization.

@@

Another track taken for the function \texttt{laned} is to unroll the loop and hash on separate lanes, and then mix the lanes together upon exiting the loop. Each lane has its own dependency chain, but on less iterations.
\\\\
\textbf{Benchmark}\\
Here are the timing on both an x86 and an ARM CPU. It also includes timing for the function \texttt{unrolled}, to show that performance increase comes indeed from ILP and not the loop unrolling itself.
We can see that \texttt{temp} and \texttt{laned} performed equally, leveraging ILP for a significant performance increase over the \texttt{baseline}.
\\
\begin{figure}[h]
    \centering  % This command centers the table 
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        CPU & \texttt{baseline} & \texttt{unrolled} & \texttt{temp} & \texttt{laned} \\
        \hline
        AMD Ryzen 5 5625U (x86 64-bit) & 92.787 µs & 93.047 µs & 37.516 µs & 37.434 µs \\
        Apple M1 Pro (ARM 64-bit) & 125.23 µs & 124.42 µs & 28.507 µs & 30.716 µs \\
        \hline
    \end{tabular}
    \caption{Benchmark timings for ILP example}
    \label{tab:your_table_label}
\end{figure}
\\
While the \texttt{temp} and \texttt{laned} functions won't yield exactly the same hashes as the \texttt{baseline}, they serve the same purpose, while being much faster. The following section will delve more in depth onto the definition of such an high ILP construction, for performance critical scenarios.

\begin{figure}[h]
    \begin{multicols}{2}
        \begin{lstlisting}[language=Rust, style=boxed]
const PRIME: u64 = 0x00000100000001b3;
const OFFSET: u64 = 0xcbf29ce484222325;

@@

\subsection{Laned Construction}

In order to improve ILP, a construction can processes the message by groups of \( k_b \) blocks. This way, each group is hashed independently and intermediate hashes are compressed altogether thereafter.

@@

Let's define \( n_g = \lfloor {n_b}/{k_b} \rfloor \) as the number of whole groups of \( k_b \) message blocks. \\
For each group we compute an intermediate hash, \( H_i \), as follows:

@@

\section{GxHash Algorithm}

>>>>

\documentclass[10pt]{article} 
\usepackage[margin=0.8in]{geometry}
\usepackage{amsmath}
\usepackage{hyperref} 
\usepackage{listings}
\usepackage{xcolor}
\usepackage{multicol}

\input{rust-listings.tex}

@@

\subsection{The Merkle–Damgård Construction}

@@

\subsection{The Wide-Pipe Construction} \label{widepipe}

@@

While compilers and CPUs employ various techniques to optimize ILP, their capabilities are often constrained by the inherent data dependencies
in the code. It's for this reason that algorithmic design can be pivotal. By structuring algorithms to minimize dependency chains from the outset,
we create opportunities for higher ILP that even the most advanced compiler optimizations and CPU features cannot achieve alone. Therefore,
algorithmic design that is mindful of ILP can be a game-changer for performance optimization.

\subsubsection{Example}
Let's take a FNV-like hashing function. The "naive" way to process an array of elements would look like the \texttt{baseline} method,
as shown in Rust snippet (Figure~\ref{fig:ilp_rust_example}). As we can see, every loop iteration requires the value of \texttt{h},
which has been computed the iteration before. Here we have a dependency chain, preventing the compiler from doing any optimization.

@@

Another track taken for the function \texttt{laned} is to unroll the loop and hash on separate lanes,
and then mix the lanes together upon exiting the loop. Each lane has its own dependency chain, but on less iterations.

\begin{figure}[ht]
\begin{multicols}{2}
\begin{lstlisting}[language=Rust, style=boxed]
const PRIME: u64 = 0x00000100000001b3;
const OFFSET: u64 = 0xcbf29ce484222325;

@@

\clearpage
\subsubsection{Benchmark}
Here are the timing on both an x86 and an ARM CPU. It also includes timing for the function \texttt{unrolled},
to show that performance increase comes indeed from ILP and not the loop unrolling itself.
We can see that \texttt{temp} and \texttt{laned} performed equally, leveraging ILP for a significant performance increase over the \texttt{baseline}.
\\
\begin{figure}[ht]
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
CPU & \texttt{baseline} & \texttt{unrolled} & \texttt{temp} & \texttt{laned} \\
\hline
AMD Ryzen 5 5625U (x86 64-bit) & 92.787 µs & 93.047 µs & 37.516 µs & 37.434 µs \\
Apple M1 Pro (ARM 64-bit) & 125.23 µs & 124.42 µs & 28.507 µs & 30.716 µs \\
\hline
\end{tabular}
\caption{Benchmark timings for ILP example}
\label{tab:your_table_label}
\end{figure}
\\
While the \texttt{temp} and \texttt{laned} functions won't yield exactly the same hashes as the \texttt{baseline},
they serve the same purpose, while being much faster.
Both approaches have they pro and cons. As the \texttt{laned} explicitely declares \( n \) variables for our lanes,
this approach is simpler in regards to compiler analysis and is thus more likely to benefit from ILP,
regardless of the compiler or programming language used.
The following section will delve more in depth onto the definition of a \textbf{Laned Construction}.

\clearpage
\subsection{The Laned Construction}

The \textbf{Laned Construction} introduces \( k_b \) lanes, and processes the message by groups of \( k_b \) blocks,
so that for a given group each of the \( k_b \) block and be processed on its own lane.
This way, we have \( k_b \) independent dependency chains. The lane hashes can then be compressed altogether thereafter upon exiting the loop.

@@

Let's define \( n_g = \lfloor {n_b}/{k_b} \rfloor \) as the number of whole groups of \( k_b \) message blocks. \\
For each lane we compute an intermediate hash, \( H_i \), as follows:

@@

\section{The GxHash Algorithm}
